---
title: "Shark Attacks - Perception, Sentiment, and Data"
author: "Tim Hogan"
date: "12/16/2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
library(leaflet)
library(leaflet.extras)
library(tidyverse)
library(magrittr)
library(readxl)
library(tigris)
library(downloader)
library(stringr)
library(tidytext)
library(wordcloud)
library(reshape2)
library(gridExtra)
library(rvest)

#Read in data and subset rows and columns that aren't NA's. Remove cases that are missing.
attack <- read_xls("GSAF5-2.xls")
attack <- attack[1:6483,1:16]
attack <- attack[-which(is.na(attack$`Case Number...1`)),]

#Change year to number. Create new variables to quantify aspects.
#Hurt is if there was actually an injury. GSAF is if it was reported by GSAF first.
#common is if it's Provoked or Unprovoked. Confirmed is if shark characteristics had been recorded.
attack$Year <- as.numeric(attack$Year)
attack$Hurt <- ifelse(grepl("No injury",attack$Injury)==T,0,1)
attack$GSAF <- ifelse(grepl("GSAF",attack$`Investigator or Source`)==T,1,0)
attack$common <- ifelse(attack$Type=="Unprovoked"|attack$Type=="Provoked",1,0)
attack$Confirmed <- ifelse(grepl("No shark",attack$Species)==T | is.na(attack$Species)==T|grepl("involvement",attack$Species)==T,0,1)
attack$`Fatal (Y/N)`[is.na(attack$`Fatal (Y/N)`)] <- 0
attack %<>% separate(Date,c("Day","Month","Yearr"),sep="-")

#Set year boundary past 2010. Change everything else to factors for analysis.
attack2010 <- subset(attack,Year>=2010 & Country=="USA")
attack2010$Hurt <- as.factor(attack2010$Hurt)
attack2010$common <- as.factor(attack2010$common)
attack2010$Confirmed <- as.factor(attack2010$Confirmed)
attack2010$Laceration <- ifelse(grepl("aceratio",attack2010$Injury)==T,1,0)

# Pull in states for leaflet.
states <- states(cb=T)

source("TextFunctions.R")

textdf <- readtext("https://www.trackingsharks.com/mans-inflatable-paddleboard-sunk-by-10ft-tiger-shark-shark-hits-kid/",
                   "https://www.trackingsharks.com/great-white-bites-kayak-in-california-leaves-2-massive-teeth/",
                   3,3); textdf$set <- "2019-Unharmed-TC"
textdf1 <- readtext("https://www.trackingsharks.com/shark-bite-snorkelers-arm-down-to-bone/",
                    "https://www.trackingsharks.com/shark-bites-surfer-near-cocoa-beach-pier/",
                    3,3); textdf1$set <- "2019-Harmed-TC"
textdf2 <- readtext("https://www.hawaiinewsnow.com/2019/06/15/teen-loses-leg-fingers-shark-attack/",
                    "https://www.hawaiinewsnow.com/2019/08/20/shark-attack-reported-napoopoo-beach-park-off-hawaii-island/",
                    1,2); textdf2$set <- "2019-Harmed-HNN"

textdf3 <- readtext("https://www.hawaiinewsnow.com/story/20010243/shark-attack-survivor-recounts-frightening-ordeal/",
                    "https://www.hawaiinewsnow.com/story/20230172/beaches-closed-after-man-by-shark-on-maui/",
                    1,2); textdf3$set <- "2010-Harmed-HNN"
```

## Map of Shark Attacks

```{r,echo=F}
sumattack <- attack2010 %>% subset(common==1) %>% group_by(Area) %>% summarize(Attacks=n(),Provoked=sum(Type=="Provoked"),Uninjured=sum(Hurt==0), Fatal=sum(`Fatal (Y/N)`=="Y"))

states_merged <- geo_join(states,sumattack,"NAME","Area")
states_merged <- subset(states_merged, !is.na(Attacks))

pal <- colorNumeric("Reds",domain=states_merged@data[,c("Attacks")])
#states_merged$n[is.na(states_merged$n)] <- 0

popup_attack <- paste0("Total: ", as.character(states_merged$n),
                       "\n Provoked: ", as.character(states_merged$Provoked),
                       "\n Uninjured: ", as.character(states_merged$Uninjured))

leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%
  setView(-108.483330,38.712046,zoom=3) %>%
  addPolygons(data=states_merged,
              fillColor=~pal(states_merged@data[,c("Attacks")]),
              fillOpacity=0.7,
              weight=0.2,
              smoothFactor=0.2,
              popup=~popup_attack) %>%
  addLegend(pal=pal,
            values=states_merged@data[,c("Attacks")],
            position="bottomright",
            title="Shark Attacks")

```

## How do people feel about sharks?

- Big teeth
- Sharp teeth
- Jaws theme
- Bad movies
- Big teeth

## However, they're important, and endangered.

- Many are predators that control the food chain effectively.
- As some become endangered, these ecosystems can fall out of balance.
- This can cause degradation and a reduction in important species to us.
- Conservationists have been working to preserve sharks, but have to work against this negative image.

## Why the bad image?

- Jaws is the mass-media start of sharks being scary ocean monsters.
- Shows like Shark Week have romanticized shark attack stories.
- Overall, sharks are mostly associated with shark attacks.

## Questions

- How frequent and bad are shark attacks?
- Also, how do journalism articles cover shark attacks?
- The Global Shark Attack File has records of shark attacks and links to sources.
- This project will focus on the United States, between 2010 and 2019.

## Number of Attacks

```{r,echo=F}
ggplot(attack2010) + geom_bar(aes(x=Year,fill=Type)) + xlab("") + ylab("Frequency")  + scale_x_continuous(breaks=seq(2010,2019,1))
```

- There are also some shark "interactions", not just "attacks".

## Fatalities by State

```{r,echo=F}
pal1 <- colorNumeric("Reds",domain=states_merged@data[,c("Fatal")])
#states_merged$n[is.na(states_merged$n)] <- 0

popup_attack <- paste0("Total: ", as.character(states_merged$n),
                       "\n Provoked: ", as.character(states_merged$Provoked),
                       "\n Uninjured: ", as.character(states_merged$Uninjured))

leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%
  setView(-108.483330,38.712046,zoom=3) %>%
  addPolygons(data=states_merged,
              fillColor=~pal1(states_merged@data[,c("Fatal")]),
              fillOpacity=0.7,
              weight=0.2,
              smoothFactor=0.2,
              popup=~popup_attack) %>%
  addLegend(pal=pal1,
            values=states_merged@data[,c("Fatal")],
            position="bottomright",
            title="Fatalities")
```

## Map colored by injuries

```{r,echo=F}
pal2 <- colorNumeric("Reds",domain=states_merged@data[,c("Uninjured")])
#states_merged$n[is.na(states_merged$n)] <- 0

popup_attack <- paste0("Total: ", as.character(states_merged$n),
                       "\n Provoked: ", as.character(states_merged$Provoked),
                       "\n Uninjured: ", as.character(states_merged$Uninjured))

leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%
  setView(-108.483330,38.712046,zoom=3) %>%
  addPolygons(data=states_merged,
              fillColor=~pal2(states_merged@data[,c("Uninjured")]),
              fillOpacity=0.7,
              weight=0.2,
              smoothFactor=0.2,
              popup=~popup_attack) %>%
  addLegend(pal=pal2,
            values=states_merged@data[,c("Uninjured")],
            position="bottomright",
            title="Fatalities")
```

## Infrequent, but still not the best

```{r,echo=F}
ggplot(attack2010) + geom_bar(aes(x=Year,fill=as.factor(Laceration))) + xlab("") + ylab("Frequency") + scale_x_continuous(breaks=seq(2010,2019,1)) + labs(fill="Laceration")
```

- Lacerations are scary and can happen pretty frequently.

## Media Analysis

- In this presentation, I'll compare two groups of articles.
- Two are about bad attacks in 2019 and two are about bad attacks in 2010.
- They're short, so I'll analyze them in pairs.

## Sentiment comparison: 2010 and 2019.

```{r,echo=F,message=F}
source("TextFunctions.R")

p1 <- sentorder(textdf3) + theme(legend.position="none") + labs(title="2010",subtitle="From Hawaii News Now")
p2 <- sentorder(textdf2) + theme(legend.position="none") + labs(title="2019",subtitle="From Hawaii News Now")
grid.arrange(p1,p2,ncol=2)

```

## However, let's look closer.

```{r,echo=F,message=F}
  snrc <- get_sentiments("nrc")
  sent <- inner_join(textdf2, snrc)
  topsent <- sent %>% group_by(word,sentiment) %>% summarize(n=n()) %>% arrange(-n)
  ggplot(data=topsent[1:20,],aes(x=word,y=n,fill=sentiment)) + geom_bar(stat="identity") + coord_flip() + 
  ylab("Number of Occurrences") + xlab("") + theme(legend.position = "none") + theme_bw() + labs(title="Word frequency (and sentiment) in analyzed 2019 articles")

```

## Compare to Conservation Group.
```{r,echo=F,message=F}
  snrc <- get_sentiments("nrc")
  sent <- inner_join(textdf, snrc)
  topsent <- sent %>% group_by(word,sentiment) %>% summarize(n=n()) %>% arrange(-n)
  ggplot(data=topsent[1:20,],aes(x=word,y=n,fill=sentiment)) + geom_bar(stat="identity") + coord_flip() + 
  ylab("Number of Occurrences") + xlab("") + theme(legend.position = "none") + theme_bw() + labs(title="Word frequency (and sentiment) in 2019 articles with no injuries",subtitle="Articles from TrackingSharks.com")

```

## What does this mean?

- Shark attacks aren't frequent, but a lot of the injuries are still bad.
- When covering a shark attack, the focus is on the injury, which makes the shark seem scary.
- The conservation article seems to be even more focused on the injury than the news.

## What can be done next?

- To improve a shark's image, covering them outside of shark attacks may be the key.
- Analyzing more shark articles not focused on shark attacks can provide an interesting point of comparison.
- Additionally, analyzing and comparing these factors to countries like Australia may provide a lot of insight into other nation-specific factors of these effects.

# Thank you!



